---
title: "Data quality check"
output: html_document
author: "Anil and Steve"
---

# Boilerplate

```{r setup, include=FALSE}
rm(list=ls())
options(warn=-1)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r}
# library(reasontheme)
# library(pensionviewr)
#library(janitor)
library(tidyverse)
library(plyr)
#library(openxlsx)
library(tseries)
#library(ggplot2)
library(data.table)
library(openxlsx)
#library(readr)
library(rsconnect)
library(base64enc)
#Shiny-----------
library(shiny)
# library(shinyWidgets)
# library(shinymanager)
# library(repmis)
#library(shinyFiles)
library(DT)
library(plotly)
library(grid)
library(readr)
#Shiny
library(rlang)
library(purrr)
library(rpart)
# library(vip)
# library(skimr)
# library(renv)
# library(profvis)
library(nortest)
library(car)
```

# pullStateData() function

```{r}
pullStateData <- function (FY)
{
  con <- RPostgres::dbConnect(
    RPostgres::Postgres(),
    dbname = "d629vjn37pbl3l",
    host = "ec2-3-209-200-73.compute-1.amazonaws.com",
    port = 5432,
    user = "reason_readonly",
    password = "p88088bd28ea68027ee96c65996f7ea3b56db0e27d7c9928c05edc6c23ef2bc27",
    sslmode = "require")
  
  query <- paste("select * from pull_data_state_only()\nwhere year > '",
                 paste(FY - 1), "'\nand attribute_name in ('1 Year Investment Return Percentage',
                 \n'1-Yr Investment Return',
                 \n'PPD ID',
                 \n'Number of Employers',
                 \n'Number of Plans',
                 \n'Actuarial Cost Method in GASB Reporting',
                 \n'Employer Normal Cost Percentage',
                 \n'Total Normal Cost Dollar',
                 \n'Actuarial Funded Ratio Percentage',
                 \n'Market Investment Return (MVA Basis)',
                 \n'Actuarial Value of Assets GASB Dollar',
                 \n'Market Value of Assets Dollar',
                 \n'Market Assets Reported for Asset Smoothing',
                 \n'Actuarially Accrued Liabilities Dollar',
                 \n'Total Pension Liability Dollar',
                 \n'Actuarially Determined Contribution Percentage of Payroll',
                 \n'Actuarially Required Contribution Dollar',
                 \n'Actuarially Required Contribution Paid Percentage',
                 \n'Total Employer Contributions',
                 \n'Statutory Payment Dollar',
                 \n'Statutory Payment Percentage',
                 \n'Amortizaton Method',
                 \n'Total Benefits Paid Dollar',
                 \n'Benefit Payments Dollar',
                 \n'Refunds Dollar',
                 \n'Administrative Expense Dollar',
                 \n'Cost Structure',
                 \n'Asset Valuation Method for GASB Reporting',
                 \n'Covered Payroll Dollar',
                 \n'Employee Contribution Dollar',
                 \n'Employee Normal Cost Percentage',
                 \n'Employer Contribution Regular Dollar',
                 \n'Employer Normal Cost Dollar',
                 \n'Employer State Contribution Dollar',
                 \n'ADEC as a Percent of Payroll',
                 \n'Other Contribution Dollar',
                 \n'Other Additions',
                 \n'Fiscal Year of Contribution',
                 \n'Inflation Rate Assumption For GASB Reporting',
                 \n'Investment Return Assumption for GASB Reporting',
                 \n'Discount Rate Assumption',
                 \n'Number of Years Remaining on Amortization Schedule',
                 \n'Payroll Growth Assumption',
                 \n'Unfunded Actuarially Accrued Liabilities Dollar',
                 \n'Total Contribution Dollar',
                 \n'Total Normal Cost Percentage',
                 \n'Net Pension Liability Dollar',
                 \n'Net Pension Liability Assuming 1% Decrease in Discount Rate',
                 \n'Net Pension Liability Assuming 1% Increase in Discount Rate',
                 \n'Total Number Of Members',
                 \n'Total Projected Actuarial Required Contribution Percentage Of Payroll',
                 \n'Actuarially Required Contribution Paid Dollar',
                 \n'Actuarially Determined Contribution Paid Dollar',
                 \n'Type of Employees Covered',
                 \n'Total Amortization Payment Percentage',
                 \n'Wage Inflation')")
  result <- RPostgres::dbSendQuery(con, query)
  all_data <- RPostgres::dbFetch(result) %>% janitor::clean_names()
  RPostgres::dbClearResult(result)
  RPostgres::dbDisconnect(con)
  all_data %>% dplyr::group_by_at(dplyr::vars(-.data$attribute_value)) %>%
    dplyr::mutate(row_id = 1:dplyr::n()) %>% dplyr::ungroup() %>%
    tidyr::pivot_wider(names_from = attribute_name, values_from = attribute_value) %>%
    dplyr::select(-.data$row_id) %>% 
    dplyr::arrange(display_name, year) %>%
    janitor::clean_names()
}
```

# filterData2() functions

```{r}
filterData2 <- function(Data, fy = 2001,
                        inflation = TRUE,
                        expenditure = TRUE,
                        employee = NULL,
                        sp500 = TRUE,
                        dcData = TRUE,
                        treasury = TRUE,
                        blend.teacher = FALSE,
                        source = FALSE) {
  if (isTRUE(blend.teacher)) {
    blended.teach <- c("Arizona State Retirement System",
                       "Delaware State Employees’ Pension Plan", "District of Columbia Teachers Retirement Fund",
                       "Florida Retirement System", "Employee Retirement System of Hawaii",
                       "Idaho Public Employee Retirement System", "Iowa Public Employees' Retirement System",
                       "Kansas Public Employees' Retirement System", "Public Employees' Retirement System of Mississippi",
                       "Nevada Public Employees Retirement System", "New Hampshire Retirement System",
                       "North Carolina Teachers' and State Employees' Retirement System",
                       "Oregon Public Employees Retirement System", "Rhode Island Employees Retirement System",
                       "South Carolina Retirement Systems", "South Dakota Retirement System",
                       "Tennessee Consolidated Retirement System, Teachers Pension Plan",
                       "Utah Retirement Systems, Noncontributory Retirement System",
                       "Virginia Retirement System", "Wisconsin Retirement System",
                       "Wyoming Retirement System, Public Employees’ Pension Plan",
                       "Maine Public Employees Retirement System (PERS) Defined Benefit Plan",
                       "Maryland State Employees’ Retirement System")
    Data <- data.frame(Data)
    Data <- Data %>% mutate(type_of_employees_covered = case_when(display_name %in%
                                                                    blended.teach ~ "Plan covers state, local and teachers",
                                                                  TRUE ~ type_of_employees_covered))
  }
  Data <- data.frame(Data)
  columns <- c("total_pension_liability_dollar", "wage_inflation",
               "payroll_growth_assumption", "payroll_growth_assumption_ppd", "other_contribution_dollar",
               "other_additions_dollar", "x1_year_investment_return_percentage",
               "amortizaton_method", "number_of_years_remaining_on_amortization_schedule",
               "amortization_payment_total_amount",
               "x1_year_investment_return_percentage",
               "x1_year_investment_return_percentage_ppd",
               "total_normal_cost_dollar", "fiscal_year_of_contribution",
               "statutory_payment_dollar", "statutory_payment_percentage",
               "ppd_fund_id", "number_of_employers", "number_of_plans", "ppd_id",
               "discount_rate_assumption", "market_investment_return_mva_basis", 
               "actuarially_determined_contribution_dollar",
               "net_pension_liability_dollar",
               "administrative_expense_dollar",
               "employer_state_contribution_dollar",
               "total_amortization_payment_percentage",
               "investment_return_assumption_for_gasb_reporting",
               "type_of_employees_covered",
               "total_contribution_dollar",
               "total_normal_cost_percentage_ppd",
               "employer_contribution_regular_dollar",
               "actuarial_cost_method_in_gasb_reporting",
               "net_pension_liability_assuming_1_percent_decrease_in_discount_rate",
               "net_pension_liability_assuming_1_percent_increase_in_discount_rate",
               "actuarially_required_contribution_paid_dollar",
               "actuarially_required_contribution_paid_percentage",
               "total_benefits_paid_dollar",
               "benefit_payments_dollar",
               "refunds_dollar",
               "actuarially_determined_contribution_paid_dollar",
               "cost_structure", "employer_normal_cost_percentage", "actuarial_value_of_assets_dollar",
               "asset_valuation_method_for_gasb_reporting", "inflation_rate_assumption",
               "total_number_of_members", "total_projected_actuarial_required_contribution_percentage_of_payroll",
               "market_assets_reported_for_asset_smoothing")
  columns2 <- c("test")
  for (i in (1:length(columns))) {
    if (sum(colnames(Data) %in% columns[i]) == 0) {
      columns2 <- rbind(columns2, columns[i])
    }
  }
  columns2 <- as.character(columns2 <- columns2[2:length(columns2)])
  cols <- matrix(NA, length(Data[, 1]), length(columns2))
  colnames(cols) <- columns2
  Data <- cbind(Data, cols)
  Data <- data.frame(Data)
  Data <- Data %>% arrange(state, display_name, year)
  Data$year <- as.numeric(Data$year)
  if (isTRUE(inflation)) {
    library(readxl)
    library(httr)
    url1 <- "https://github.com/ReasonFoundation/databaseR/raw/master/files/CPI_by_Region_DOL.xlsx"
    GET(url1, write_disk(tf <- tempfile(fileext = ".xlsx")))
    cpi <- read_excel(tf, sheet = 1)
    regions <- read_excel(tf, sheet = 2)
    cpi <- data.table(cpi)
    regions <- data.table(regions)
    cpi$cpi_change <- NA
    for (i in (2:length(cpi$region))) {
      cpi$cpi_change[i] <- (cpi$cpi[i]/cpi$cpi[i - 1]) -
        1
    }
    Data <- Data %>% left_join(regions, by = "state")
    Data <- Data %>% left_join(cpi, by = c("year", "region"))
    Data <- Data %>% select(-cpi)
  }
  
  if (isTRUE(sp500)) {
    urlfile2 = "https://raw.githubusercontent.com/ReasonFoundation/databaseR/master/files/S%26P500.csv"
    
    
    SP500 <- data.table(
      read_csv(url(urlfile2), col_names = TRUE, na = c(""), 
               skip_empty_rows = TRUE, col_types = NULL))
    
    Data <- Data %>% left_join(SP500, by = c("year"))
    
  }
  
  if (isTRUE(dcData)) {
    
    #Data <- data.table(pullStateData(2001))
    
    urlfile3 = "https://raw.githubusercontent.com/ReasonFoundation/databaseR/master/files/DCPlanData.csv"
    
    DCPlanData <- data.table(
      read_csv(url(urlfile3), col_names = TRUE, na = c(""), 
               skip_empty_rows = TRUE, col_types = NULL))
    
    DCPlanData <- DCPlanData %>% 
      select(year = fy, ppd_id, plantype,
             ee_contrate_mandatory, er_contrate_fixedrate, 
             planname, primary,
             eligible_eegroups, assets, members_total, actives, inactives, coveredpayroll) %>% 
      replace(is.na(.), 0)
    
    DCPlanData$ppd_id <- as.numeric(DCPlanData$ppd_id)
    DCPlanData$fy <- as.numeric(DCPlanData$fy)
    DCPlanData$er_contrate_fixedrate <- as.numeric(DCPlanData$er_contrate_fixedrate)
    DCPlanData$ee_contrate_mandatory <- as.numeric(DCPlanData$ee_contrate_mandatory)
    DCPlanData$assets <- as.numeric(DCPlanData$assets)
    DCPlanData$members_total <- as.numeric(DCPlanData$members_total)
    DCPlanData$actives <- as.numeric(DCPlanData$actives)
    DCPlanData$inactives <- as.numeric(DCPlanData$inactives)
    
    for(i in (1:length(DCPlanData$members_total))){
      
      DCPlanData$members_total[i] <- max(DCPlanData$members_total[i], (DCPlanData$actives[i]+DCPlanData$inactives[i]))
      
    }
    
    Data$ppd_id <- as.numeric(Data$ppd_id)
    Data$year<- as.numeric(Data$year)
    
    Data <- Data %>% left_join(DCPlanData, by = c("ppd_id", "year"))
    
  }
  if (isTRUE(treasury)) {
    urlfile <- "https://raw.githubusercontent.com/ReasonFoundation/databaseR/master/files/treasury.csv"
    treasury <- data.table(
      read_csv(url(urlfile), col_names = TRUE, na = c(""), skip_empty_rows = TRUE, col_types = NULL))
    
    colnames(treasury)[2] <- "treasury_30"
    Data <- Data %>% left_join(treasury, by = c("year"))
    
  }
  
  if (isTRUE(expenditure)) {
    
    #1. General Fund Exp./Rev. NASBO (State)
    #2. Total Exp./Rev. NASBO (State) --> general funds, other state funds, bonds, and federal funds.
    #3. Total Exp./Rev. Urban (State & Local) --> state and local direct general expenditures (exclude “business-like” activities such as utilities and transfers between state and local governments.)
    
    library(readxl)
    library(httr)
    urlfile="https://raw.githubusercontent.com/ReasonFoundation/databaseR/master/files/NASBO_Exp_Report_Data_v1.csv"
    NASBO<-read_csv(url(urlfile), col_names = TRUE, na = c(""), skip_empty_rows = TRUE, col_types = NULL)
    NASBO <- setDT(NASBO)
    NASBO <- NASBO %>% select(YEAR, STATE,"TOTAL_CAPI") %>% filter(YEAR>2000)
    #View(NASBO)
    colnames(NASBO)[1] <- "year"
    colnames(NASBO)[2] <- "state"
    Data <- Data %>% left_join(NASBO, by = c("year","state"))
    ####
    urlfile="https://raw.githubusercontent.com/ReasonFoundation/databaseR/master/files/UrbanCensus_ExpRev_2000-2018.csv"
    Urban<-read_csv(url(urlfile), col_names = TRUE, na = c(""), skip_empty_rows = TRUE, col_types = NULL)
    Urban <- setDT(Urban)
    
    Urban <- Urban %>% select(Year, State,"(R01) Total Revenue", "(E001) Total Expenditure")
    
    colnames(Urban)[1] <- "year"
    colnames(Urban)[2] <- "state"
    
    Data <- Data %>% left_join(Urban, by = c("year","state"))
    
  }
  
  Data <- Data %>% select(year, plan_name = display_name,
                          state, if (isTRUE(source)) {
                            "data_source_name"
                          }, return_1yr = x1_year_investment_return_percentage,return_1yr_ppd = x1_year_investment_return_percentage_ppd,
                          ava_return = market_investment_return_mva_basis, actuarial_cost_method_in_gasb_reporting,
                          funded_ratio = actuarial_funded_ratio_percentage, ava = actuarial_value_of_assets_gasb_dollar,
                          mva = market_value_of_assets_dollar, #mva_smooth = market_assets_reported_for_asset_smoothing,
                          aal = actuarially_accrued_liabilities_dollar,
                          adec1 = actuarially_determined_contribution_dollar,
                          adec2 =   actuarially_required_contribution_paid_dollar,  #actuarially_required_contribution_dollar (PPD)
                          adec_paid_pct = actuarially_required_contribution_paid_percentage,
                          adec_pct = adec_as_a_percent_of_payroll, #instead of employers_projected_actuarial_required_contribution
                          arc_paid_dollar = actuarially_required_contribution_paid_dollar, #NEW
                          adec_paid_dollar = actuarially_determined_contribution_paid_dollar,#NEW
                          statutory = statutory_payment_dollar, statutory_pct = statutory_payment_percentage,
                          amortizaton_method, total_benefit_payments = total_benefits_paid_dollar,
                          benefit_payments = benefit_payments_dollar, refunds = refunds_dollar,
                          admin_exp = administrative_expense_dollar, cost_structure,
                          state_exp_NASBO = TOTAL_CAPI, #gf_rev_NASBO = "TOTAL REV",
                          state.local_exp_Census = "(R01) Total Revenue", state.local_rev_Census = "(E001) Total Expenditure",
                          fund_id = ppd_fund_id, ppd_id = ppd_id, employers = number_of_employers,
                          plans = number_of_plans, asset_valuation_method_for_gasb_reporting,
                          dc_plantype = plantype,  dc_er_pct = er_contrate_fixedrate, #dc_eligble = eligible_eegroups,
                          dc_ee_pct = ee_contrate_mandatory, dc_plan = planname, dc_primary = primary,
                          dc_assets = assets, dc_members = members_total, dc_payroll = coveredpayroll,
                          payroll = covered_payroll_dollar, ee_contribution = employee_contribution_dollar,
                          ee_nc_pct = employee_normal_cost_percentage,
                          er_nc_pct = employer_normal_cost_percentage,
                          total_nc_pct = total_normal_cost_percentage, 
                          total_nc_pct_ppd = total_normal_cost_percentage_ppd, 
                          er_contribution = employer_contribution_regular_dollar, er_state_contribution = employer_state_contribution_dollar,
                          other_contribution = other_contribution_dollar, other_additions = other_additions_dollar,
                          fy_contribution = fiscal_year_of_contribution, inflation_assum = inflation_rate_assumption,
                          arr = investment_return_assumption_for_gasb_reporting,
                          dr = discount_rate_assumption, number_of_years_remaining_on_amortization_schedule,
                          payroll_growth_assumption,payroll_growth_assumption_ppd, total_amortization_payment_pct = total_amortization_payment_percentage,
                          total_amo_payment = amortization_payment_total_amount, #Added
                          total_contribution = total_contribution_dollar,
                          total_nc_dollar = total_normal_cost_dollar, total_number_of_members,
                          total_proj_adec_pct = total_projected_actuarial_required_contribution_percentage_of_payroll,
                          tpl = total_pension_liability_dollar,
                          npl_pct_down = net_pension_liability_assuming_1_percent_decrease_in_discount_rate, #Added
                          npl = net_pension_liability_dollar, #Added
                          npl_pct_up = net_pension_liability_assuming_1_percent_increase_in_discount_rate, #Added
                          type_of_employees_covered, unfunded_actuarially_accrued_liabilities_dollar,
                          wage_inflation, region, cpi_change, sp_500, treasury_30)
  
  
  ## Combining ARC & ADEC
  Data$state.local_exp_Census <- as.numeric(Data$state.local_exp_Census)
  Data$state.local_rev_Census <- as.numeric(Data$state.local_exp_Census)
  Data$state_exp_NASBO <- as.numeric(Data$state_exp_NASBO)
  Data$dc_primary <- ifelse(Data$dc_primary == 1, "Yes", "No")
  
  Data$adec1 <- as.numeric(Data$adec1)
  Data$adec2 <- as.numeric(Data$adec2)
  Data$adec_paid_dollar <- as.numeric(Data$adec_paid_dollar)
  Data$arc_paid_dollar <- as.numeric(Data$arc_paid_dollar)
  Data$arc_adec_dollar <- ifelse(is.na(Data$adec1) == TRUE,Data$adec2, Data$adec1)
  Data$arc_adec_paid_dollar <- ifelse(is.na(Data$arc_paid_dollar) == TRUE,
                                      ifelse(is.na(Data$statutory_payment_dollar) == TRUE,statutory_payment_dollar, 
                                             Data$adec_paid_dollar), 
                                      Data$arc_paid_dollar)
  
  Data$fy_contribution <- as.numeric(Data$fy_contribution)
  Data$year <- as.numeric(Data$year)
  Data$fy_contribution <- round(Data$fy_contribution, 0)
  Data <- Data %>% filter(year >= fy) #%>% select(-adec1, -adec2)
  
  #https://www.urban.org/policy-centers/cross-center-initiatives/state-and-local-finance-initiative/projects/state-fiscal-briefs/florida
  
  if (isTRUE(expenditure)) {
    Data$ava <- as.numeric(Data$ava)
    Data$aal <- as.numeric(Data$aal)
    Data$state_exp_NASBO <- as.numeric(Data$state_exp_NASBO)
    Data$state.local_exp_Census <- as.numeric(Data$state.local_exp_Census)
    Data <- Data %>% group_by(state, year) %>% mutate(ual_exp_NASBO = round((sum(aal)-sum(ava))/(state_exp_NASBO*1000000),2),
                                                      ual_exp_Census = round((sum(aal)-sum(ava))/(state.local_exp_Census*1000),2))
  }
  
  if (is_null(employee)) {
    employee <- employee
  }
  else if (employee == "teacher") {
    employee <- c("Plan covers teachers")
  }
  else if (employee == "state and local") {
    employee <- c("Plan covers state and local employees")
  }
  else if (employee == "police and fire") {
    employee <- c("Plan covers police and/or fire")
  }
  else if (employee == "state") {
    employee <- c("Plan covers state employees")
  }
  else if (employee == "local") {
    employee <- c("Plan covers local employees")
  }
  else if (employee == "state, local, and teachers") {
    employee <- c("Plan covers state, local and teachers")
  }
  if (is_null(employee)) {
    Data <- data.frame(Data) %>% replace(is.na(.), 0)
  }
  else {
    Data %>% filter(type_of_employees_covered == paste(employee)) %>% replace(is.na(.), 0)
  }
}

```


```{r, match plan names}

#### [1] Load PPD data ####
## Need tidyverse and dplyr packages
# urlfile="https://raw.githubusercontent.com/ReasonFoundation/databaseR/master/files/ppd-data-2022-04-01.csv"
# PPD<-read_csv(url(urlfile), col_names = TRUE, na = c(""), skip_empty_rows = TRUE, col_types = NULL)
# PPD <- setDT(PPD)

#### [2] Load Reason data ####
# reason.data <- pullStateData(2001)
# reason.data <- filterData2(reason.data, 2001)

#### [3] Save unique/distinct ppd ids (Reason database) ####
# reason.id <- unique(reason.data$ppd_id)

#### [4] Find the same Reason dataabse ids in PPD database ####
# View(PPD %>% filter(ppd_id %in% reason.id))

### Example with ppd_id == 10
# View(data.frame(
#   PPD %>% filter(ppd_id == 10) %>% select(ppd.name = PlanName),
#   reason.data %>% filter(ppd_id == 10) %>% select(reason.name = plan_name)))

#########
```

# Check data quality
## Notes

Project components:
* Compare Reason pension data with ppd data
(Variables/Years where Reason and PPD data is different)
     * Key variables (aal, mva, ava, inv.return, NC %, contribution amount)
     * All state-level plans (aggregate and plan-by-plan)
     * Across multiple periods (2001-21, 2/5/10-year)
     * Unpaired t-tests (same/diff variance, normal distribution)
* Improve the data importing/collecting process at Reason
     * Develop easy to implement/general data test for data collectors (automated)

All of tests does not point out the significantly incorrect value of aal in 2019


## Data
### Update w/ latest PPD file

```{r}
urlfile="https://raw.githubusercontent.com/ReasonFoundation/databaseR/master/files/ppd-data-Q2-2022.csv"
PPD <- read_csv(url(urlfile), col_names = TRUE, na = c(""), skip_empty_rows = TRUE, 
                col_types = NULL)
PPD <- data.table(PPD)
PPD <- PPD %>% filter(AdministeringGovt == 0)
```

```{r}
## Turn all values numeric
PPD$ActAssets_AVA  <- as.numeric(PPD$ActAssets_AVA) * 1000
PPD$MktAssets_net  <- as.numeric(PPD$MktAssets_net ) * 1000
PPD$ActFundedRatio_GASB  <- as.numeric(PPD$ActFundedRatio_GASB)
PPD$ActLiabilities_GASB  <- as.numeric(PPD$ActLiabilities_GASB) * 1000
PPD$payroll  <- as.numeric(PPD$payroll) * 1000
PPD$InvestmentReturn_1yr <- as.numeric(PPD$InvestmentReturn_1yr)
PPD$contrib_tot <- as.numeric(PPD$contrib_tot) * 1000
```

```{r}
PPD <- PPD[(2001 <= PPD$fy) & (PPD$fy <= 2021),]
```


### Pull reason data

```{r}
reason.data <- pullStateData(2001)
reason.data <- filterData2(reason.data, 2001)
reason.data <- data.table(reason.data)
```

```{r}
## Turn all values numeric
reason.data$mva <- as.numeric(reason.data$mva)
reason.data$ava <- as.numeric(reason.data$ava)
reason.data$funded_ratio <- as.numeric(reason.data$funded_ratio)
reason.data$aal <- as.numeric(reason.data$aal)
reason.data$payroll <- as.numeric(reason.data$payroll)
reason.data$total_contribution <- as.numeric(reason.data$total_contribution)
reason.data$return_1yr <- as.numeric(reason.data$return_1yr)
```

```{r}
reason.data <- reason.data[(2001 <= reason.data$year) & (reason.data$year <= 2021), ]
```


```{r}
year_start <- 2001
year_end <- 2021
```


### Extract state-level data

```{r}
reason_plan_names <- reason.data %>% select(plan_name, ppd_id) %>% distinct()
PPD_plan_names <- PPD %>% select(PlanName, ppd_id) %>% distinct()
merged_plan_names <- merge(x=reason_plan_names, y=PPD_plan_names, 
                           by="ppd_id", all.x=T)
colnames(merged_plan_names) <- c("ppd_id", "reason_names", "PPD_names")
```

```{r}
# three plans are missed in PPD
merged_plan_names[merged_plan_names$ppd_id == 0]
```

## Statistical tests
### Normality tests
#### Test functions

```{r}
# Shapiro-Wilk test
sw_test <- function(data){
  return(shapiro.test(data))
}
```


```{r}
# Anderson-Darling Test
ad_test <- function(data){
  return(ad.test(reason.data$aal))
}
```

```{r}
normality_test_func <- function(stat_test, datasets, variable_names){
  normality_tests <- matrix(0, nrow = length(reason_variable_names), ncol = 2)
  rownames(normality_tests) <- reason_variable_names
  colnames(normality_tests) <- c('Reason data', 'PPD data')
  
  for(col in 1:2){
    for(row in 1:7){
      data <- na.omit(datasets[[col]][[variable_names[[col]][row]]])
      if(length(data) <= 3){
        normality_tests[row, col] <- "NA"
      }
      else{
        normality_tests[row, col] <- stat_test(data)$p.value
      }
    }
  }
  return(normality_tests)
}
```

#### Normality test outcomes

```{r}
reason_variable_names <- c("mva", "ava", "funded_ratio", "aal", "payroll", "total_contribution", "return_1yr")
PPD_variable_names <- c("MktAssets_net", "ActAssets_AVA", "ActFundedRatio_GASB", "ActLiabilities_GASB",
                        "payroll", "contrib_tot", "InvestmentReturn_1yr")
lst_variable_names <- list(reason_variable_names, PPD_variable_names)
```

**Normality tests for aggregate data**

```{r}
datasets <- list(reason.data, PPD)
normality_test_func(stat_test = sw_test, datasets = datasets, variable_names = lst_variable_names)
```

**Normality tests for each plan**

```{r}
indPlan_normalTest <- function(ppd_id, stat_test, variable_names){
  reason_data <- reason.data %>%
      filter(plan_name == merged_plan_names[merged_plan_names$ppd_id == ppd_id]$reason_names) %>%
      select(variable_names[[1]])
  PPD_data <- PPD %>%
      filter(PlanName == merged_plan_names[merged_plan_names$ppd_id == 1]$PPD_names) %>%
      select(variable_names[[2]])
  datasets <- list(reason_data, PPD_data)
  return(normality_test_func(stat_test = sw_test, datasets = datasets, variable_names = lst_variable_names))
}
```


```{r}
indPlan_normalTest(ppd_id=100, stat_test = sw_test, variable_names = lst_variable_names)
```

In addition to the statistic tests, we should visually check the distribution. In theory, p-value measures the probability of observing the empirical data given that the null hypothesis is true. Thus, any deviation from a normal distribution also make p-value large. This is illustrated by the plot below


```{r}
plot_distribution <- function(dataset, variable_name){
  data <- dataset[[variable_name]]
  mean_ <- mean(data, na.rm=T)
  sd_ <- sd(data, na.rm=T)
  den <- density(data, na.rm=T)
  plot(NULL, xlim=c(min(den$x), max(den$x)), ylim=c(min(den$y), max(den$y)),
            xlab="value", ylab="density", main=paste0(variable_name, "'s distribution"))
  lines(den, col="red")
  for(i in 1:100){
    lines(density(rnorm(1000, mean_, sd_)), add=T, col=alpha("gray", 0.2))
  }
}
```

```{r}
plot_distribution(dataset = reason.data, variable_name = lst_variable_names[[1]][7])
```


Another visual way to check the normality of a data is the quantile-quantile plot

In theory, if the points form a straight line, the data should come from a normal distribution

```{r}
qq_plot <- function(dataset, variable_name){
  return(qqPlot(dataset[[variable_name]]))
}
```

```{r}
qq_plot(dataset=reason.data, variable_name = lst_variable_names[[1]][7])
```


All three methods, statistic tests, QQ-plots, density plots give us information about the distribution of each variable.

We should leverage all threes to make conclusions before moving to the next steps


```{r}
plot_distribution(dataset = PPD, variable_name = lst_variable_names[[2]][7])
```

```{r}
qq_plot(dataset = PPD, variable_name = lst_variable_names[[2]][7])
```


### Unpaired tests
#### Test functions

```{r}
# unpaired t_test
unpaired_t_test <- function(reason_data_col, ppd_data_col){
  return(t.test(na.omit(reason_data_col), na.omit(ppd_data_col), var.equal=T, paired=F))
}
```

```{r}
# unpaired welch t_test
unpaired_welch_t_test <- function(reason_data_col, ppd_data_col){
  return(t.test(na.omit(reason_data_col), na.omit(ppd_data_col), var.equal=F, paired = F))
}
```

```{r}
# unpaired wilcox_test
unpaired_wilcox_test <- function(reason_data_col, ppd_data_col){
  return(wilcox.test(na.omit(reason_data_col), na.omit(ppd_data_col), alternative = "two.sided", paired = F))
}
```

#### Unpaired test outcomes

```{r}
aggregate_plan_tests <- function(stat_test, variable_names, variable_idx){
  reason_variable_name <- lst_variable_names[[1]][variable_idx]
  PPD_variable_name <- lst_variable_names[[2]][variable_idx]
  return(stat_test(na.omit(reason.data[[reason_variable_name]]), na.omit(PPD[[PPD_variable_name]])))
}
```

```{r}
aggregate_plan_tests(stat_test = unpaired_t_test, variable_names = lst_variable_names, variable_idx = 7)
```


```{r}
individual_plan_tests <- function(ppd_id, stat_test, variable_names, variable_idx){
  reason_data <- reason.data %>%
      filter(plan_name == merged_plan_names[merged_plan_names$ppd_id == ppd_id]$reason_names) %>%
      select(variable_names[[1]])
  PPD_data <- PPD %>%
      filter(PlanName == merged_plan_names[merged_plan_names$ppd_id == 1]$PPD_names) %>%
      select(variable_names[[2]])
  reason_variable_name <- lst_variable_names[[1]][variable_idx]
  PPD_variable_name <- lst_variable_names[[2]][variable_idx]
  return(stat_test(na.omit(reason_data[[reason_variable_name]]), na.omit(PPD_data[[PPD_variable_name]])))
}
```

```{r}
individual_plan_tests(ppd_id = 1, stat_test = unpaired_t_test,
                      variable_names = lst_variable_names, variable_idx = 7)
```

#### Paired tests

```{r}
# paired t_test
paired_t_test_aal <- function(reason_data_col, ppd_data_col){
  return(t.test(reason_data_col, ppd_data_col, var.equal=T, paired=T))
}
```

```{r}
# paired welch t_test
paired_welch_t_test_aal <- function(reason_data_col, ppd_data_col){
  return(t.test(reason_data_col,ppd_data_col, var.equal=FALSE, paired = T))
}
```

```{r}
# paired wilcox_test
paired_wilcox_test_aal <- function(reason_data_col, ppd_data_col){
  return(wilcox.test(reason_data_col, ppd_data_col, alternative = "two.sided", paired = T))
}
```

# Other data checks





```{r}
```

